---
title: "Schedule C Funding of Arts and Cultural Organizations"
format: 
  html:
    code-fold: true
execute:
  echo: false
---

```{r}
#| label: setup
#| include: false
#| warning: false
#| message: false

# Use quarto render to generate a static site in /docs
# Set github pages to deploy from main branch, /docs

# Insure packages are installed by checking required packages against installed packages
reqd_pkgs <- c("flextable",
               "sf",
               "tidyverse",
               "mapgl",
               "htmltools",
               "officer",
               "scales")

install.packages(
  setdiff(
    reqd_pkgs,
    rownames(installed.packages())
  ),
  repos = "https://cloud.r-project.org"
)

lapply(reqd_pkgs, library, character.only = TRUE) |> 
  invisible()

# If needed, ensure that the Quarto render and RStudio read the same environment vars
# readRenviron(".Renviron")

set_flextable_defaults(
  font.family  = "Authentic Sans",
  font.size    = 10,        # ~13px (flextable uses points)
  border.color = "#0b0b0b",
  background.color = "transparent",
  padding.top    = 8,
  padding.bottom = 4,
  padding.left   = 0,
  padding.right  = 16
)

# Wraps a data frame in a CIL-styled flextable ready to embed in the page.
# Columns size to their content by default. `widths` overrides that: one value
# per column, or named values setting just those columns. `max_width` caps the
# fitted widths, never the ones `widths` names.
# Usage: data |> cil_table(widths = c("Purpose of Funds" = 4))
cil_table <- function(data, widths = NULL, heights = NULL,
                      max_width = NULL, unit = "in") {
  ft <- flextable(data)
  fitted <- !(is.null(widths) && is.null(max_width))

  if (!fitted) {
    # Browser-side content fitting, like Excel's Autofit. width = 0 leaves the
    # table unconstrained; width = 1 would stretch it to 100% regardless of text.
    ft <- set_table_properties(ft, width = 0, layout = "autofit")
  } else {
    # Fitted widths are the baseline, capped by max_width, then overridden by
    # anything `widths` names. "autofit" discards widths, so this must be fixed.
    fit <- unlist(dim_pretty(ft)$widths)
    if (!is.null(max_width)) fit <- pmin(fit, max_width)
    if (is.null(widths)) {
      widths <- fit
    } else if (!is.null(names(widths)) && all(nzchar(names(widths)))) {
      unknown <- setdiff(names(widths), names(data))
      if (length(unknown)) {
        stop("`widths` names no such column: ", paste(unknown, collapse = ", "),
             call. = FALSE)
      }
      fit[match(names(widths), names(data))] <- widths
      widths <- fit
    } else if (length(widths) != ncol(data)) {
      # Otherwise they land on the first columns, misaligning every one.
      stop("`widths` has ", length(widths), " value(s) but `data` has ",
           ncol(data), " column(s). Give one width per column, or name them.",
           call. = FALSE)
    }
    ft <- set_table_properties(ft, layout = "fixed")
    ft <- width(ft, j = seq_along(widths), width = widths, unit = unit)
  }

  if (!is.null(heights)) {
    # Heights are only emitted under the "exact" rule; "auto" ignores them.
    ft <- height(ft, height = heights, part = "body", unit = unit)
    ft <- hrule(ft, rule = "exact", part = "body")
  }

  ft |>
    border_remove() |>
    bold(bold = TRUE, part = "header") |>
    hline(
      border = fp_border(color = "#0b0b0b", width = 0.75),
      part   = "all"
    ) |>
    padding(
      padding.top = 0, padding.bottom = 4,
      padding.left = 0, padding.right = 16,
      part = "header"
    ) |>
    # Both steps are needed to keep the table out of a shadow DOM, which no page
    # CSS can reach: ft.shadow drops the static host, and the rename hides the
    # wrapper from tabwid.js, which shadow-roots every .tabwid on DOMContentLoaded.
    htmltools_value(ft.shadow = FALSE) |>
    as.character() |>
    sub(pattern = 'class="tabwid"', replacement = 'class="cil-tabwid"', fixed = TRUE) |>
    HTML() |>
    div(class = if (fitted) "scroll-output scroll-output--fit" else "scroll-output")
}

```

```{r}
#| label: fetch-schedule-c
#| include: false
#| warning: false
#| message: false

# Version Schedule C data by release date
# Commit under data/raw
# The version is pinned by last modified date

FUNDED_DISCLOSURE_URL <- "https://rnd.council.nyc.gov/src/assets/excelFund/funded_disclosure_FY2027.xlsx"
FUNDED_DISCLOSURE_VERSION <- "2026-07-02"

raw_dir <- file.path("data", "raw")
manifest <- file.path("data", "raw-mainifest.csv")
dir.create(raw_dir, recursive = TRUE, showWarnings = FALSE)

# Get publication date by server response, or NA if not returned
upstream_version <- function(url) {
  headers <- try(curlGetHeaders(url), silent = TRUE)
  if (inherits(headers, "try-error")) return(NA_character_)

  stamp <- grep("^[Ll]ast-[Mm]odified:", headers, value = TRUE)[1]
  if (is.na(stamp)) return(NA_character_)

  parsed <- strptime(sub("^[^:]+:[[:space:]]*", "", stamp),
                     "%a, %d %b %Y %H:%M:%S", tz = "GMT")
  format(as.Date(parsed))
}


# Path to `version` of `url`, downloading it if we don't already have it. Every
# fetch is logged to the manifest, so each version's checksum is recorded rather
# than merely asserted.
fetch_raw <- function(url, version, name = basename(url)) {
  versioned <- sub("\\.([^.]+)$", sprintf("_%s.\\1", version), name)
  dest      <- file.path(raw_dir, versioned)

  if (!file.exists(dest)) {
    available <- upstream_version(url)
    if (!identical(available, version)) {
      stop("Requested version ", version, " of ", name, " is not on disk, and ",
           "the server is now serving ", available %||% "an unknown version",
           ". Recover it from git history, or bump the pin.", call. = FALSE)
    }

    message("Downloading ", versioned)
    # Write to a temp file first
    part <- paste0(dest, ".part")
    download.file(url, part, mode = "wb", quiet = TRUE)
    file.rename(part, dest)

    write.table(
      data.frame(file = versioned, version = version,
                 md5 = unname(tools::md5sum(dest)),
                 fetched = format(Sys.Date()), url = url),
      manifest, sep = ",", row.names = FALSE,
      col.names = !file.exists(manifest), append = file.exists(manifest)
    )
  }

  dest
}

funded_disclosure_path <- fetch_raw(FUNDED_DISCLOSURE_URL, FUNDED_DISCLOSURE_VERSION)

# Surface newer releases as information, not as a problem to fix.
local({
  available <- upstream_version(FUNDED_DISCLOSURE_URL)
  if (!is.na(available) && available > FUNDED_DISCLOSURE_VERSION) {
    message("A newer disclosure file is available (", available, "). ",
            "Bump FUNDED_DISCLOSURE_VERSION to use it.")
  }
})

schedule_c <- readxl::read_xlsx(funded_disclosure_path)
```

New York City Council discretionary funding to non-profits and agency initiatives (colloquially known as *Schedule C* funding) is a large source of funding for cultural and arts organizations in New York City. Getting a complete view of Schedule C funding of the sector (or any single sector, by extension) is not straightforward as:

- expense and capital funding disclosure data are released by fiscal year, via Excel files on the Council's website, and 
- awards are not classified or categorized beyond their `Source`, the initiative or provision from which the funds are awarded.

Take, for example, the first 25 awards of FY 2027 by award number `MOCS ID#`:

```{r}
cil_table(
  schedule_c |> head(25),
  widths = c("MOCS ID#" = 1.5,"Legal Name" = 3, "Purpose of Funds" = 9, "Amount" = 0.8)
)
```

There are two immediate ways to identify awards to arts and cultural organizations:

1. The initiative or provision in `Source` explicitly references arts or culture
2. The contracting city agency is the Department of Cultural Affairs (`Agency`: `DCLA`)

There are `r nrow(schedule_c |> distinct(Source))` funding initiatives for FY 2027. Those that explicitly apply to arts and cultural organizations are:

* Coalition Theaters of Color
* Cultural After-School Adventure (CASA)
* Cultural Immigrant Initiative
* Film Accelerator Model
* SU-CASA
* Support for Arts Instruction
* Community Composting at Cultural Institutions Groups
* Creative Arts Team

```{r}
arts_culture_inits <- c(
  "Coalition Theaters of Color",
  "Cultural After-School Adventure (CASA)",
  "Cultural Immigrant Initiative",
  "Film Accelerator Model",
  "SU-CASA",
  "Support for Arts Instruction",
  "Community Composting at Cultural Institutions Groups",
  "Creative Arts Team"
)

arts_culture_awards <- schedule_c |> 
  filter(Source %in% arts_culture_inits | Agency == "DCLA")
```

Filtering for the above two conditions returns `r scales::comma(nrow(arts_culture_awards))` of `r scales::comma(nrow(schedule_c))` awards, a total of `r round((nrow(arts_culture_awards)/nrow(schedule_c)*100),1)`%. These awards comprise only `r round(sum(as.numeric(arts_culture_awards$Amount))/sum(as.numeric(schedule_c$Amount))*100,1)`% of total Schedule C funds, `r scales::dollar(sum(as.numeric(arts_culture_awards$Amount)))` out of `r scales::dollar(sum(as.numeric(schedule_c$Amount)))`

This approach is limited by the omission of organizations doing arts or cultural work, but whose Schedule C funding is not easily identifiable as such. 



