---
title: "Schedule C Funding of Arts and Cultural Organizations"
format: 
  html:
    code-fold: true
---

```{r}
#| label: setup
#| include: false
#| warning: false
#| message: false

# Use quarto render to generate a static site in /docs
# Set github pages to deploy from main branch, /docs

# Insure packages are installed by checking required packages against installed packages
reqd_pkgs <- c("flextable",
               "sf",
               "tidyverse",
               "mapgl",
               "htmltools",
               "officer")

install.packages(
  setdiff(
    reqd_pkgs,
    rownames(installed.packages())
  ),
  repos = "https://cloud.r-project.org"
)

lapply(reqd_pkgs, library, character.only = TRUE) |> 
  invisible()

# If needed, ensure that the Quarto render and RStudio read the same environment vars
# readRenviron(".Renviron")

set_flextable_defaults(
  font.family  = "Authentic Sans",
  font.size    = 10,        # ~13px (flextable uses points)
  border.color = "#0b0b0b",
  background.color = "transparent",
  padding.top    = 8,
  padding.bottom = 4,
  padding.left   = 0,
  padding.right  = 16
)

# Wraps a data frame in a CIL-styled flextable ready to embed in the page.
# Usage: data |> cil_table()
cil_table <- function(data) {
  data |>
    flextable() |>
    set_table_properties(layout = "fixed") |>
    border_remove() |>
    bold(bold = TRUE, part = "header") |>
    hline(
      border = fp_border(color = "#0b0b0b", width = 0.75),
      part   = "all"
    ) |>
    padding(
      padding.top = 0, padding.bottom = 4,
      padding.left = 0, padding.right = 16,
      part = "header"
    ) |>
    htmltools_value() |>
    div(class = "scroll-output")
}

```

New York City Council discretionary funding to non-profits and agency initiatives (colloquially known as *Schedule C* funding) is a significant source of funding for cultural and arts organizations in New York City. Getting a complete view of Schedule C funding of the sector (or any single sector, by extension) is not straightforward:

- Expense and capital funding disclosure data are released annually, by fiscal year, via Excel files on the Council's website
- Awards are not classified or categorized in any way.

Take, for example, the first 10 awards of FY 2027 by award number `MOCS ID#`:

```{r}
#| label: fetch-schedule-c
#| include: false
#| warning: false
#| message: false

# Version Schedule C data by release date
# Commit under data/raw
# The version is pinned by last modified date

FUNDED_DISCLOSURE_URL <- "https://rnd.council.nyc.gov/src/assets/excelFund/funded_disclosure_FY2027.xlsx"
FUNDED_DISCLOSURE_VERSION <- "2026-07-02"

raw_dir <- file.path("data", "raw")
manifest <- file.path("data", "raw-mainifest.csv")
dir.create(raw_dir, recursive = TRUE, showWarnings = FALSE)

# Get publication date by server response, or NA if not returned
upstream_version <- function(url) {
  headers <- try(curlGetHeaders(url), silent = TRUE)
  if (inherits(headers, "try-error")) return(NA_character_)

  stamp <- grep("^[Ll]ast-[Mm]odified:", headers, value = TRUE)[1]
  if (is.na(stamp)) return(NA_character_)

  parsed <- strptime(sub("^[^:]+:[[:space:]]*", "", stamp),
                     "%a, %d %b %Y %H:%M:%S", tz = "GMT")
  format(as.Date(parsed))
}


# Path to `version` of `url`, downloading it if we don't already have it. Every
# fetch is logged to the manifest, so each version's checksum is recorded rather
# than merely asserted.
fetch_raw <- function(url, version, name = basename(url)) {
  versioned <- sub("\\.([^.]+)$", sprintf("_%s.\\1", version), name)
  dest      <- file.path(raw_dir, versioned)

  if (!file.exists(dest)) {
    available <- upstream_version(url)
    if (!identical(available, version)) {
      stop("Requested version ", version, " of ", name, " is not on disk, and ",
           "the server is now serving ", available %||% "an unknown version",
           ". Recover it from git history, or bump the pin.", call. = FALSE)
    }

    message("Downloading ", versioned)
    # Write to a temp file first
    part <- paste0(dest, ".part")
    download.file(url, part, mode = "wb", quiet = TRUE)
    file.rename(part, dest)

    write.table(
      data.frame(file = versioned, version = version,
                 md5 = unname(tools::md5sum(dest)),
                 fetched = format(Sys.Date()), url = url),
      manifest, sep = ",", row.names = FALSE,
      col.names = !file.exists(manifest), append = file.exists(manifest)
    )
  }

  dest
}

funded_disclosure_path <- fetch_raw(FUNDED_DISCLOSURE_URL, FUNDED_DISCLOSURE_VERSION)

# Surface newer releases as information, not as a problem to fix.
local({
  available <- upstream_version(FUNDED_DISCLOSURE_URL)
  if (!is.na(available) && available > FUNDED_DISCLOSURE_VERSION) {
    message("A newer disclosure file is available (", available, "). ",
            "Bump FUNDED_DISCLOSURE_VERSION to use it.")
  }
})

schedule_c <- readxl::read_xlsx(funded_disclosure_path)
```



```{r}
cil_table(schedule_c |> 
            head(10))
```
