---
title: "Schedule C Funding of Arts and Cultural Organizations"
format: 
  html:
    code-fold: true
execute:
  echo: false
---

```{r}
#| label: setup
#| include: false
#| warning: false
#| message: false

# Use quarto render to generate a static site in /docs
# Set github pages to deploy from main branch, /docs

# Insure packages are installed by checking required packages against installed packages
reqd_pkgs <- c("flextable",
               #"sf",
               "tidyverse",
               #"mapgl",
               "htmltools",
               "officer",
               "scales",
               "nccsdata")

install.packages(
  setdiff(
    reqd_pkgs,
    rownames(installed.packages())
  ),
  repos = "https://cloud.r-project.org"
)

lapply(reqd_pkgs, library, character.only = TRUE) |> 
  invisible()

# If needed, ensure that the Quarto render and RStudio read the same environment vars
# readRenviron(".Renviron")

set_flextable_defaults(
  font.family  = "Authentic Sans",
  font.size    = 10,        # ~13px (flextable uses points)
  border.color = "#0b0b0b",
  background.color = "transparent",
  padding.top    = 8,
  padding.bottom = 4,
  padding.left   = 0,
  padding.right  = 16
)

# Wraps a data frame in a CIL-styled flextable ready to embed in the page.
# Columns size to their content by default. `widths` overrides that: one value
# per column, or named values setting just those columns. `max_width` caps the
# fitted widths, never the ones `widths` names.
# Usage: data |> cil_table(widths = c("Purpose of Funds" = 4))
cil_table <- function(data, widths = NULL, heights = NULL,
                      max_width = NULL, unit = "in") {
  ft <- flextable(data)
  fitted <- !(is.null(widths) && is.null(max_width))

  if (!fitted) {
    # Browser-side content fitting, like Excel's Autofit. width = 0 leaves the
    # table unconstrained; width = 1 would stretch it to 100% regardless of text.
    ft <- set_table_properties(ft, width = 0, layout = "autofit")
  } else {
    # Fitted widths are the baseline, capped by max_width, then overridden by
    # anything `widths` names. "autofit" discards widths, so this must be fixed.
    fit <- unlist(dim_pretty(ft)$widths)
    if (!is.null(max_width)) fit <- pmin(fit, max_width)
    if (is.null(widths)) {
      widths <- fit
    } else if (!is.null(names(widths)) && all(nzchar(names(widths)))) {
      unknown <- setdiff(names(widths), names(data))
      if (length(unknown)) {
        stop("`widths` names no such column: ", paste(unknown, collapse = ", "),
             call. = FALSE)
      }
      fit[match(names(widths), names(data))] <- widths
      widths <- fit
    } else if (length(widths) != ncol(data)) {
      # Otherwise they land on the first columns, misaligning every one.
      stop("`widths` has ", length(widths), " value(s) but `data` has ",
           ncol(data), " column(s). Give one width per column, or name them.",
           call. = FALSE)
    }
    ft <- set_table_properties(ft, layout = "fixed")
    ft <- width(ft, j = seq_along(widths), width = widths, unit = unit)
  }

  if (!is.null(heights)) {
    # Heights are only emitted under the "exact" rule; "auto" ignores them.
    ft <- height(ft, height = heights, part = "body", unit = unit)
    ft <- hrule(ft, rule = "exact", part = "body")
  }

  ft |>
    border_remove() |>
    bold(bold = TRUE, part = "header") |>
    hline(
      border = fp_border(color = "#0b0b0b", width = 0.75),
      part   = "all"
    ) |>
    padding(
      padding.top = 0, padding.bottom = 4,
      padding.left = 0, padding.right = 16,
      part = "header"
    ) |>
    # Both steps are needed to keep the table out of a shadow DOM, which no page
    # CSS can reach: ft.shadow drops the static host, and the rename hides the
    # wrapper from tabwid.js, which shadow-roots every .tabwid on DOMContentLoaded.
    htmltools_value(ft.shadow = FALSE) |>
    as.character() |>
    sub(pattern = 'class="tabwid"', replacement = 'class="cil-tabwid"', fixed = TRUE) |>
    HTML() |>
    div(class = if (fitted) "scroll-output scroll-output--fit" else "scroll-output")
}

```

```{r}
#| label: fetch-schedule-c
#| include: false
#| warning: false
#| message: false

# Version Schedule C data by release date
# Commit under data/raw
# The version is pinned by last modified date

FUNDED_DISCLOSURE_URL <- "https://rnd.council.nyc.gov/src/assets/excelFund/funded_disclosure_FY2027.xlsx"
FUNDED_DISCLOSURE_VERSION <- "2026-07-02"

raw_dir <- file.path("data", "raw")
manifest <- file.path("data", "raw-mainifest.csv")
dir.create(raw_dir, recursive = TRUE, showWarnings = FALSE)

# Get publication date by server response, or NA if not returned
upstream_version <- function(url) {
  headers <- try(curlGetHeaders(url), silent = TRUE)
  if (inherits(headers, "try-error")) return(NA_character_)

  stamp <- grep("^[Ll]ast-[Mm]odified:", headers, value = TRUE)[1]
  if (is.na(stamp)) return(NA_character_)

  parsed <- strptime(sub("^[^:]+:[[:space:]]*", "", stamp),
                     "%a, %d %b %Y %H:%M:%S", tz = "GMT")
  format(as.Date(parsed))
}


# Path to `version` of `url`, downloading it if we don't already have it. Every
# fetch is logged to the manifest, so each version's checksum is recorded rather
# than merely asserted.
fetch_raw <- function(url, version, name = basename(url)) {
  versioned <- sub("\\.([^.]+)$", sprintf("_%s.\\1", version), name)
  dest      <- file.path(raw_dir, versioned)

  if (!file.exists(dest)) {
    available <- upstream_version(url)
    if (!identical(available, version)) {
      stop("Requested version ", version, " of ", name, " is not on disk, and ",
           "the server is now serving ", available %||% "an unknown version",
           ". Recover it from git history, or bump the pin.", call. = FALSE)
    }

    message("Downloading ", versioned)
    # Write to a temp file first
    part <- paste0(dest, ".part")
    download.file(url, part, mode = "wb", quiet = TRUE)
    file.rename(part, dest)

    write.table(
      data.frame(file = versioned, version = version,
                 md5 = unname(tools::md5sum(dest)),
                 fetched = format(Sys.Date()), url = url),
      manifest, sep = ",", row.names = FALSE,
      col.names = !file.exists(manifest), append = file.exists(manifest)
    )
  }

  dest
}

funded_disclosure_path <- fetch_raw(FUNDED_DISCLOSURE_URL, FUNDED_DISCLOSURE_VERSION)

# Surface newer releases as information, not as a problem to fix.
local({
  available <- upstream_version(FUNDED_DISCLOSURE_URL)
  if (!is.na(available) && available > FUNDED_DISCLOSURE_VERSION) {
    message("A newer disclosure file is available (", available, "). ",
            "Bump FUNDED_DISCLOSURE_VERSION to use it.")
  }
})

schedule_c <- readxl::read_xlsx(funded_disclosure_path)
```

New York City Council discretionary funding to non-profits and agency initiatives (colloquially known as *Schedule C* funding) is a large source of funding for cultural and arts organizations in New York City. Getting a complete view of Schedule C funding of the sector (or any single sector, by extension) is not straightforward as:

- expense and capital funding disclosure data are released by fiscal year, via Excel files on the Council's website, and 
- awards are not classified or categorized beyond their `Source`, the initiative or provision from which the funds are awarded.

Take, for example, the first 25 awards of FY 2027 by award number `MOCS ID#`:

```{r}
cil_table(
  schedule_c |> head(25),
  widths = c("MOCS ID#" = 1.5,"Legal Name" = 3, "Purpose of Funds" = 9, "Amount" = 0.8)
)
```

There are two immediate ways to identify awards to arts and cultural organizations:

1. The initiative or provision in `Source` explicitly references arts or culture
2. The contracting city agency is the Department of Cultural Affairs (`Agency`: `DCLA`)

There are `r nrow(schedule_c |> distinct(Source))` funding initiatives for FY 2027. Those that explicitly apply to arts and cultural organizations are:

* Coalition Theaters of Color
* Cultural After-School Adventure (CASA)
* Cultural Immigrant Initiative
* Film Accelerator Model
* SU-CASA
* Support for Arts Instruction
* Community Composting at Cultural Institutions Groups
* Creative Arts Team

```{r}
arts_culture_inits <- c(
  "Coalition Theaters of Color",
  "Cultural After-School Adventure (CASA)",
  "Cultural Immigrant Initiative",
  "Film Accelerator Model",
  "SU-CASA",
  "Support for Arts Instruction",
  "Community Composting at Cultural Institutions Groups",
  "Creative Arts Team"
)

arts_culture_awards <- schedule_c |> 
  filter(Source %in% arts_culture_inits)
```

Filtering for the above arts and culture funding initiatives returns `r scales::comma(nrow(arts_culture_awards))` of `r scales::comma(nrow(schedule_c))` awards, a total of `r round((nrow(arts_culture_awards)/nrow(schedule_c)*100),1)`%. These awards comprise only `r round(sum(as.numeric(arts_culture_awards$Amount))/sum(as.numeric(schedule_c$Amount))*100,1)`% of total Schedule C funds, `r scales::dollar(sum(as.numeric(arts_culture_awards$Amount)))` out of `r scales::dollar(sum(as.numeric(schedule_c$Amount)))`

This approach is limited by the omission of organizations whose Schedule C award(s) funding source do not easily indicate their engagement in arts or cultural work.

Another pathway to Schedule C award classification is to use Federal IRS data. The IRS classifies nonprofit organizations using the National Taxonomy of Exempt Entities (NTEE) system, which was developed by the National Center For Charitable Statistics in the 1980s. Nonprofits are assigned an NTEE code during the tax-exempt application process. The NCCS outlines appropriate uses of NTEE codes as follows:

> *The National Taxonomy of Exempt Entities (NTEE) is a descriptive taxonomy that helps regulators and researchers aggregate data for statistical analysis. While certain nonprofit characteristics like the assigned 501(c) subsection code can have significant binding implications such as whether a donation can be treated as tax-deductible (501c3 public charities) or NOT (most other 501c types). The NTEE classification, on the other hand, should not be used as a limiting factor by funders or other institutional actors. The NTEE is often not precise and many nonprofits have missions that span multiple NTEE categories. It is NOT recommended to ever eliminate nonprofits from funding considerations because of their NTEE code or assume that their NTEE category captures their full range of activities.*

Nonprofit NTEE codes are included in the IRS' Business Master File (BMF) data. BMF data may be joined to the Schedule C disclosure records via Federal Tax ID numbers (EINs). 

```{r}
# Get NCCS BMF cols
dict <- nccs_dictionary()

# Format NYCC EIN strings to NCCS standard
schedule_c <- schedule_c |> 
  mutate(ein_nccs = nccs_normalize_ein(`Tax ID`))

# Get lazy arrow query to append with dplyr chain before collecting
query <- nccs_read(collect = FALSE)

# Pull down BMF rows for orgs in FY27 Schedule C dataset
result <- query |> 
  filter(ein %in% schedule_c$ein_nccs) |> 
  collect()

# Join BMF to Schedule C on EIN
schedule_c_bmf <- schedule_c |> 
  left_join(result, by = join_by("ein_nccs" == "ein"))

# Unique EINs in Schedule C funding
unique_ein <- schedule_c_bmf |> distinct(ein_nccs) |> nrow()

# Successful joins
eins_joined <- schedule_c_bmf |> filter(!is.na(org_name_display)) |> distinct(ein_nccs) |> nrow()

# Unsuccessful joins
eins_notJoined <- schedule_c_bmf |> filter(is.na(org_name_display)) |> distinct(ein_nccs) |> nrow()
```

Of the `r scales::comma(unique_ein)` unique organizations awarded Schedule C funds in FY27, `r scales::comma(eins_joined)` (`r round(eins_joined/unique_ein*100,1)`%) can successfully be joined to BMF data. `r eins_notJoined` organizations cannot:

```{r}
# TODO: Format table to fill div horizontally

cil_table(schedule_c_bmf |> filter(is.na(org_name_display)) |> distinct(`Legal Name`, `Tax ID`) |> arrange(`Tax ID`))
```

Broadly, these organizations without a BMF record are:

- City entities sharing the Tax ID `136400434`
- CUNY (`133893536`)
- Houses of worship
- Others

```{r}
has_ntee <- schedule_c_bmf |> 
  filter(!is.na(org_name_display) & !(ntee_code_definition %in% c("UNDEFINED", "Unknown"))) |> 
  distinct(ein_nccs) |> 
  nrow()
```

Of the `r scales::comma(eins_joined)` organizations able to be joined to BMF data, `r scales::comma(has_ntee)` have an NTEE classification (i.e. their code is not `UNDEFINED` or `Unknown`).

<!--
From here, separate out schedule_c_bmf into two groups:

- Has NTEE classification
- Unknown NTEE code OR unable to be joined to BMF

Then, classify first group by arts and culture NTEE classification

Then, compare to the original funding source classification

Ultimately, this approach will append arts-classified orgs to the list based on funding source (i.e. no double counting)


Unused copy:

Both of these approaches are imperfect:

* Funding for arts and cultural organizations or programming is not limited to initiatives or provisions that make an explicit reference to the sector
* DCLA is not the only agency that funds arts and culture organizations
* Organizations of all types may engage in arts and cultural programming

-->