---
title: "Schedule C Funding of Arts and Cultural Organizations"
format: 
  html:
    code-fold: true
execute:
  echo: false
---

```{r}
#| label: setup
#| include: false
#| warning: false
#| message: false

# Use quarto render to generate a static site in /docs
# Set github pages to deploy from main branch, /docs

# Insure packages are installed by checking required packages against installed packages
reqd_pkgs <- c("flextable",
               #"sf",
               "tidyverse",
               #"mapgl",
               "htmltools",
               "officer",
               "scales",
               "nccsdata")

install.packages(
  setdiff(
    reqd_pkgs,
    rownames(installed.packages())
  ),
  repos = "https://cloud.r-project.org"
)

lapply(reqd_pkgs, library, character.only = TRUE) |> 
  invisible()

# If needed, ensure that the Quarto render and RStudio read the same environment vars
# readRenviron(".Renviron")

set_flextable_defaults(
  font.family  = "Authentic Sans",
  font.size    = 10,        # ~13px (flextable uses points)
  border.color = "#0b0b0b",
  background.color = "transparent",
  padding.top    = 8,
  padding.bottom = 4,
  padding.left   = 0,
  padding.right  = 16
)

# Wraps a data frame in a CIL-styled flextable ready to embed in the page.
# Columns size to their content by default. `widths` overrides that: one value
# per column, or named values setting just those columns. `max_width` caps the
# fitted widths, never the ones `widths` names.
# Usage: data |> cil_table(widths = c("Purpose of Funds" = 4))
cil_table <- function(data, widths = NULL, heights = NULL,
                      max_width = NULL, unit = "in", full_width = FALSE) {
  ft <- flextable(data)
  fitted <- !(is.null(widths) && is.null(max_width))
  
  if (fitted && full_width) {
    stop("`full_width` stretches columns to the container; `widths`/`max_width` ",
         "pin them. Use one or the other.", call. = FALSE)
  }

  if (!fitted) {
    # Browser-side content fitting, like Excel's Autofit. width = 0 leaves the
    # table unconstrained; width = 1 would stretch it to 100% regardless of text.
    ft <- set_table_properties(ft, if (full_width) 1 else 0, layout = "autofit")
  } else {
    # Fitted widths are the baseline, capped by max_width, then overridden by
    # anything `widths` names. "autofit" discards widths, so this must be fixed.
    fit <- unlist(dim_pretty(ft)$widths)
    if (!is.null(max_width)) fit <- pmin(fit, max_width)
    if (is.null(widths)) {
      widths <- fit
    } else if (!is.null(names(widths)) && all(nzchar(names(widths)))) {
      unknown <- setdiff(names(widths), names(data))
      if (length(unknown)) {
        stop("`widths` names no such column: ", paste(unknown, collapse = ", "),
             call. = FALSE)
      }
      fit[match(names(widths), names(data))] <- widths
      widths <- fit
    } else if (length(widths) != ncol(data)) {
      # Otherwise they land on the first columns, misaligning every one.
      stop("`widths` has ", length(widths), " value(s) but `data` has ",
           ncol(data), " column(s). Give one width per column, or name them.",
           call. = FALSE)
    }
    ft <- set_table_properties(ft, layout = "fixed")
    ft <- width(ft, j = seq_along(widths), width = widths, unit = unit)
  }

  if (!is.null(heights)) {
    # Heights are only emitted under the "exact" rule; "auto" ignores them.
    ft <- height(ft, height = heights, part = "body", unit = unit)
    ft <- hrule(ft, rule = "exact", part = "body")
  }

  ft |>
    border_remove() |>
    bold(bold = TRUE, part = "header") |>
    hline(border = fp_border(color = "#0b0b0b", width = 1.5),  part = "header") |>
    hline(border = fp_border(color = "#969696", width = 0.75), part = "body") |>
    padding(
      padding.top = 0, padding.bottom = 4,
      padding.left = 0, padding.right = 16,
      part = "header"
    ) |>
    # Both steps are needed to keep the table out of a shadow DOM, which no page
    # CSS can reach: ft.shadow drops the static host, and the rename hides the
    # wrapper from tabwid.js, which shadow-roots every .tabwid on DOMContentLoaded.
    htmltools_value(ft.shadow = FALSE) |>
    as.character() |>
    sub(pattern = 'class="tabwid"', replacement = 'class="cil-tabwid"', fixed = TRUE) |>
    HTML() |>
    div(class = if (fitted) "scroll-output scroll-output--fit" else "scroll-output")
}

```

```{r}
#| label: fetch-schedule-c
#| include: false
#| warning: false
#| message: false

# Version Schedule C data by release date
# Commit under data/raw
# The version is pinned by last modified date

FUNDED_DISCLOSURE_URL <- "https://rnd.council.nyc.gov/src/assets/excelFund/funded_disclosure_FY2027.xlsx"
FUNDED_DISCLOSURE_VERSION <- "2026-07-02"

raw_dir <- file.path("data", "raw")
manifest <- file.path("data", "raw-mainifest.csv")
dir.create(raw_dir, recursive = TRUE, showWarnings = FALSE)

# Get publication date by server response, or NA if not returned
upstream_version <- function(url) {
  headers <- try(curlGetHeaders(url), silent = TRUE)
  if (inherits(headers, "try-error")) return(NA_character_)

  stamp <- grep("^[Ll]ast-[Mm]odified:", headers, value = TRUE)[1]
  if (is.na(stamp)) return(NA_character_)

  parsed <- strptime(sub("^[^:]+:[[:space:]]*", "", stamp),
                     "%a, %d %b %Y %H:%M:%S", tz = "GMT")
  format(as.Date(parsed))
}


# Path to `version` of `url`, downloading it if we don't already have it. Every
# fetch is logged to the manifest, so each version's checksum is recorded rather
# than merely asserted.
fetch_raw <- function(url, version, name = basename(url)) {
  versioned <- sub("\\.([^.]+)$", sprintf("_%s.\\1", version), name)
  dest      <- file.path(raw_dir, versioned)

  if (!file.exists(dest)) {
    available <- upstream_version(url)
    if (!identical(available, version)) {
      stop("Requested version ", version, " of ", name, " is not on disk, and ",
           "the server is now serving ", available %||% "an unknown version",
           ". Recover it from git history, or bump the pin.", call. = FALSE)
    }

    message("Downloading ", versioned)
    # Write to a temp file first
    part <- paste0(dest, ".part")
    download.file(url, part, mode = "wb", quiet = TRUE)
    file.rename(part, dest)

    write.table(
      data.frame(file = versioned, version = version,
                 md5 = unname(tools::md5sum(dest)),
                 fetched = format(Sys.Date()), url = url),
      manifest, sep = ",", row.names = FALSE,
      col.names = !file.exists(manifest), append = file.exists(manifest)
    )
  }

  dest
}

funded_disclosure_path <- fetch_raw(FUNDED_DISCLOSURE_URL, FUNDED_DISCLOSURE_VERSION)

# Surface newer releases as information, not as a problem to fix.
local({
  available <- upstream_version(FUNDED_DISCLOSURE_URL)
  if (!is.na(available) && available > FUNDED_DISCLOSURE_VERSION) {
    message("A newer disclosure file is available (", available, "). ",
            "Bump FUNDED_DISCLOSURE_VERSION to use it.")
  }
})

schedule_c <- readxl::read_xlsx(funded_disclosure_path)
```
# Identifying sector funding

New York City Council discretionary funding to non-profit organizations (colloquially known as *Schedule C* funding) is a large source of funding for arts and culture in New York City. However, getting a complete view of Schedule C funding of the sector (or any single sector, by extension) is not straightforward as:

- expense and capital funding disclosure data are released by fiscal year, via Excel files on the Council's website, and 
- awards are not classified or categorized beyond their `Source`, the initiative or provision from which the funds are awarded.

Take, for example, the first 25 awards of FY 2027 by award number `MOCS ID#`:

```{r}
cil_table(
  schedule_c |> head(25),
  widths = c("MOCS ID#" = 1.5,"Legal Name" = 3, "Purpose of Funds" = 9, "Amount" = 0.8)
)
```

There are two immediate ways to identify awards to arts and cultural organizations:

1. The initiative or provision in `Source` explicitly references arts or culture
2. The contracting city agency is the Department of Cultural Affairs (`Agency`: `DCLA`)

## By Funding Initiative

There are `r nrow(schedule_c |> distinct(Source))` funding initiatives for FY 2027. Those that explicitly apply to arts and cultural organizations are:

* Coalition Theaters of Color
* Cultural After-School Adventure (CASA)
* Cultural Immigrant Initiative
* Film Accelerator Model
* SU-CASA
* Support for Arts Instruction
* Community Composting at Cultural Institutions Groups
* Creative Arts Team

```{r}
arts_culture_inits <- c(
  "Coalition Theaters of Color",
  "Cultural After-School Adventure (CASA)",
  "Cultural Immigrant Initiative",
  "Film Accelerator Model",
  "SU-CASA",
  "Support for Arts Instruction",
  "Community Composting at Cultural Institutions Groups",
  "Creative Arts Team"
)

arts_culture_awards <- schedule_c |> 
  filter(Source %in% arts_culture_inits)
```

Filtering for the above arts and culture funding initiatives returns `r scales::comma(nrow(arts_culture_awards))` of `r scales::comma(nrow(schedule_c))` awards, a total of `r round((nrow(arts_culture_awards)/nrow(schedule_c)*100),1)`%. These awards comprise only `r round(sum(as.numeric(arts_culture_awards$Amount))/sum(as.numeric(schedule_c$Amount))*100,1)`% of total Schedule C funds, `r scales::dollar(sum(as.numeric(arts_culture_awards$Amount)))` out of `r scales::dollar(sum(as.numeric(schedule_c$Amount)))`

This approach is limited by the omission of organizations whose Schedule C award(s) funding source do not easily indicate their engagement in arts or cultural work.

## By IRS BMF NTEE Code

Another pathway to Schedule C award classification is to use Federal IRS data. The IRS classifies nonprofit organizations using the National Taxonomy of Exempt Entities (NTEE) system, which was developed by the National Center For Charitable Statistics in the 1980s. Nonprofits are assigned an NTEE code during the tax-exempt application process. The NCCS outlines appropriate uses of NTEE codes as follows:

> *The National Taxonomy of Exempt Entities (NTEE) is a descriptive taxonomy that helps regulators and researchers aggregate data for statistical analysis. While certain nonprofit characteristics like the assigned 501(c) subsection code can have significant binding implications such as whether a donation can be treated as tax-deductible (501c3 public charities) or NOT (most other 501c types). The NTEE classification, on the other hand, should not be used as a limiting factor by funders or other institutional actors. The NTEE is often not precise and many nonprofits have missions that span multiple NTEE categories. It is NOT recommended to ever eliminate nonprofits from funding considerations because of their NTEE code or assume that their NTEE category captures their full range of activities.*

Nonprofit NTEE codes are included in the IRS' Business Master File (BMF) data. BMF data may be joined to the Schedule C disclosure records via Federal Tax ID numbers (EINs). 

```{r}
# Get NCCS BMF cols
dict <- nccs_dictionary()

# Format NYCC EIN strings to NCCS standard
schedule_c <- schedule_c |> 
  mutate(ein_nccs = nccs_normalize_ein(`Tax ID`))

# Get lazy arrow query to append with dplyr chain before collecting
query <- nccs_read(collect = FALSE)

# Pull down BMF rows for orgs in FY27 Schedule C dataset
result <- query |> 
  filter(ein %in% schedule_c$ein_nccs) |> 
  collect()

# Join BMF to Schedule C on EIN
schedule_c_bmf <- schedule_c |> 
  left_join(result, by = join_by("ein_nccs" == "ein"))

# Unique EINs in Schedule C funding
unique_ein <- schedule_c_bmf |> distinct(ein_nccs) |> nrow()

# Successful joins
eins_joined <- schedule_c_bmf |> filter(!is.na(org_name_display)) |> distinct(ein_nccs) |> nrow()

# Unsuccessful joins
eins_notJoined <- schedule_c_bmf |> filter(is.na(org_name_display)) |> distinct(ein_nccs) |> nrow()
```

Of the `r scales::comma(unique_ein)` unique organizations awarded Schedule C funds in FY27, `r scales::comma(eins_joined)` (`r round(eins_joined/unique_ein*100,1)`%) can successfully be joined to BMF data. `r eins_notJoined` organizations cannot:

```{r}
cil_table(schedule_c_bmf |> filter(is.na(org_name_display)) |> distinct(`Legal Name`, `Tax ID`) |> arrange(`Tax ID`),
          full_width = TRUE)
```

Broadly, these organizations without a BMF record are:

- City entities sharing the Tax ID `136400434`
- CUNY (`133893536`)
- Houses of worship
- Others

```{r}
has_ntee <- schedule_c_bmf |> 
  filter(!is.na(org_name_display) & !(ntee_code_definition %in% c("UNDEFINED", "Unknown"))) |> 
  distinct(ein_nccs) |> 
  nrow()
```

Of the `r scales::comma(eins_joined)` organizations able to be joined to BMF data, `r scales::comma(has_ntee)` have an NTEE classification (i.e. their code is not `UNDEFINED` or `Unknown`).

We can now expand the universe of arts and culture awards by including organizations whose NTEE major code is "Arts, Culture and Humanities" in addition to awards whose funding initiative is arts- or culture-based. We can also include awards with DCLA as the contracting agency. 

```{r}
addtnl_arts_awards <- schedule_c_bmf |> 
  filter(ntee_code_major_group == "Arts, Culture and Humanities" & !(`MOCS ID#` %in% arts_culture_awards$`MOCS ID#`))

arts_culture_all <- schedule_c_bmf |> 
  filter(ntee_code_major_group == "Arts, Culture and Humanities" | Source %in% arts_culture_inits | Agency == "DCLA")
```

This expands the number of arts and culture awards from `r scales::comma(nrow(arts_culture_awards))` to `r scales::comma(nrow(arts_culture_all))`. This represents `r round(nrow(arts_culture_awards)/nrow(schedule_c)*100,1)`% of FY27 Schedule C awards.

The expanded definition of arts and culture awards comprise `r round(sum(as.numeric(arts_culture_all$Amount))/sum(as.numeric(schedule_c$Amount))*100,1)`% of total Schedule C funds, `r scales::dollar(sum(as.numeric(arts_culture_all$Amount)))` out of `r scales::dollar(sum(as.numeric(schedule_c$Amount)))`.

# Top Funders of Arts & Cultural Organizations
## By Council Member

```{r}
# Spending by CM
top_cm <- arts_culture_all |> 
  group_by(`Council Member`) |> 
  summarise(Amount = sum(Amount, na.rm = T), Number = n()) |>
  mutate(pct = round(Amount / sum(Amount) * 100,3)) |> 
  arrange(desc(Amount))

# Exclude categories that aren't individual CMs
top_cm_ind <- top_cm |> 
  filter(`Council Member` != "Speaker" & str_detect(`Council Member`, "Delegation", negate = TRUE) & !is.na(`Council Member`))

top_non_cm <- top_cm |> 
  filter(!`Council Member` %in% top_cm_ind$`Council Member`)

delegation_awards <- top_cm |> 
  filter(str_detect(`Council Member`, "Delegation"))

delegation_pct <- round(sum(top_non_cm |> filter(str_detect(`Council Member`, "Delegation")) |> pull(pct)),1)
delegation_amt <- sum(top_non_cm |> filter(str_detect(`Council Member`, "Delegation")) |> pull(Amount))

cm_major_groups <- 
  bind_rows(
    # NA and Speaker
    top_cm |> filter(is.na(`Council Member`) | `Council Member` == "Speaker") |> select(`Council Member`, Amount, pct),
    # Individual CMS
    tibble(
      `Council Member` = "Individual CMs",
      Amount = sum(top_cm_ind$Amount),
      pct = sum(top_cm_ind$pct)
    ),
    # Borough Delegations
    tibble(
      `Council Member` = "Borough Delegations",
      Amount = sum(delegation_awards$Amount),
      pct = sum(delegation_awards$pct)
    )
  ) |> 
  arrange(desc(pct)) |> 
  mutate(`Council Member` = replace_na(`Council Member`, "No CM listed"))

ind_cm_pct <- cm_major_groups |> filter(`Council Member` == "Individual CMs") |> pull(pct)

ojs_define(cm_data = cm_major_groups)
```

Awards from individual Council Members to arts and cultural organizations represent the largest share of awards by `Council Member`, `r round(ind_cm_pct,1)`% (`r scales::dollar(sum(top_cm_ind$Amount))`), but not the majority.

`r round(top_cm |> head(1) |> pull(pct), 1)`% (`r scales::dollar(top_cm |> head(1) |> pull(Amount))`) of arts and culture awards do not have a Council Member associated with them (`Council Member`: `NA`) at all. `r round((top_cm |> head(2) |> pull(pct))[2],1)`% (`r scales::dollar((top_cm |> head(2) |> pull(Amount))[2])`) of awards come from the Council Speaker (`Council Member`: `Speaker`). 

The remaining `r delegation_pct`% (`r scales::dollar(delegation_amt)`) is awarded by the five Borough Delegations. 

```{ojs}
d3 = require("d3@7")
groups = transpose(cm_data)

// Based on https://gist.github.com/dbuezas/9306799
chart = {
  const width = 760, height = 440;
  const radius = Math.min(width * 0.34, height / 2 - 20);

  const svg = d3.create("svg")
      .attr("viewBox", [-width / 2, -height / 2, width, height])
      .attr("style", "max-width: 100%; height: auto; font: 13px Authentic Sans;");

  const color = d3.scaleOrdinal()
      .domain(["Individual CMs", "No CM listed", "Speaker", "Borough Delegations"])
      .range(["#3258a3", "#3e7bd2", "#73b3e7", "#c7e1f6"])
      .unknown("#8a8a8a");

  const arc      = d3.arc().innerRadius(0).outerRadius(radius * 0.8);
  const outerArc = d3.arc().innerRadius(radius * 0.9).outerRadius(radius * 0.9);
  const midAngle = d => d.startAngle + (d.endAngle - d.startAngle) / 2;
  const arcs = d3.pie().sort(null).value(d => d.pct)(groups);

  // label anchor per slice, plus a de-overlap pass the gist doesn't have
  const labels = arcs.map(d => {
    const side = midAngle(d) < Math.PI ? 1 : -1;
    return { d, side, y: outerArc.centroid(d)[1], x: radius * side };
  });

  const minGap = 34;
  for (const side of [1, -1]) {
    const col = labels.filter(l => l.side === side).sort((a, b) => a.y - b.y);
    for (let i = 1; i < col.length; i++)
      col[i].y = Math.max(col[i].y, col[i - 1].y + minGap);
    const overflow = col.length ? col[col.length - 1].y - (height / 2 - 20) : 0;
    if (overflow > 0) col.forEach(l => l.y -= overflow);
  }

  svg.append("g")
    .selectAll("path")
    .data(labels)
    .join("path")
      .attr("d", l => arc(l.d))
      .attr("fill", l => color(l.d.data["Council Member"]))
      .attr("stroke", "white")
      .attr("stroke-width", 2)
    .append("title")
      .text(l => `${l.d.data["Council Member"]}: ${l.d.data.pct.toFixed(1)}%`);

  svg.append("g")
      .attr("fill", "none")
      .attr("stroke", "#666")
      .attr("stroke-width", 1)
      .attr("opacity", 0.5)
    .selectAll("polyline")
    .data(labels)
    .join("polyline")
      .attr("points", l => [
        arc.centroid(l.d),
        outerArc.centroid(l.d),
        [l.x * 0.95, l.y]
      ]);

  svg.append("g")
    .selectAll("text")
    .data(labels)
    .join("text")
      .attr("transform", l => `translate(${l.x},${l.y})`)
      .attr("text-anchor", l => l.side === 1 ? "start" : "end")
      .attr("dy", "0.35em")
      .call(t => t.append("tspan")
          .attr("font-weight", "bold")
          .text(l => l.d.data["Council Member"]))
      .call(t => t.append("tspan")
          .attr("x", 0).attr("dy", "1.2em")
          .attr("fill", "#555")
          .text(l => `${l.d.data.pct.toFixed(1)}%`));

  return svg.node();
}
```

```{r}
top_cm <- top_cm_ind |> head(1) |> pull(`Council Member`)
top_cm_n <- top_cm_ind |> head(1) |> pull(`Number`)
top_cm_amount <- top_cm_ind |> head(1) |> pull(`Amount`)

bottom_cm <- top_cm_ind |> tail(1) |> pull(`Council Member`)
bottom_cm_n <- top_cm_ind |> tail(1) |> pull(`Number`)
bottom_cm_amount <- top_cm_ind |> tail(1) |> pull(`Amount`)
```

The amount of money awarded to arts and cultural organizations by individual council member varies widely.

Council Member `r top_cm` made `r top_cm_n` awards to arts and cultural organizations, totalling `r dollar(top_cm_amount)`.

Council Member `r bottom_cm` made only `r bottom_cm_n` awards to arts and cultural organizations, totalling `r dollar(bottom_cm_amount)`.
```{r}
top_cm_ind_reformat <- top_cm_ind |> 
  mutate(Amount = scales::dollar(Amount)) |> 
  rename(
    "Total Amount" = "Amount",
    "Number of Awards" = "Number",
    "Share of Funding (%)" = "pct"
    ) 

cil_table(top_cm_ind_reformat, full_width = TRUE)
```
## By Agency

```{r}
arts_culture_all$Agency[arts_culture_all$Agency == "dcla"] <- "DCLA"

top_agencies <- arts_culture_all |> 
  group_by(Agency) |> 
  summarise(Amount = sum(Amount, na.rm = T), Number = n()) |>
  mutate(pct = round(Amount / sum(Amount) * 100,3)) |> 
  arrange(desc(Amount))

top_agencies_grouped <- top_agencies |>
  mutate(Agency = if_else(row_number() <= 6, Agency, "All other agencies")) |>
  group_by(Agency) |>
  summarise(Amount = sum(Amount), Number = sum(Number), pct = sum(pct)) |>
  arrange(desc(Amount))

ojs_define(agency_data = top_agencies_grouped)
```

`r top_agencies |> slice(1) |> pull(Agency)` is the contracting agency for `r top_agencies |> slice(1) |> pull(pct) |> round(1)`% (`r top_agencies |> slice(1) |> pull(Amount) |> scales::dollar()`) of arts and culture awards. `r top_agencies |> slice(2) |> pull(Agency)` and `r top_agencies |> slice(3) |> pull(Agency)` are the second- and third-largest contracting agencies, with `r top_agencies |> slice(2) |> pull(pct) |> round(1)`% (`r top_agencies |> slice(2) |> pull(Amount) |> scales::dollar()`) and `r top_agencies |> slice(3) |> pull(pct) |> round(1)`% (`r top_agencies |> slice(3) |> pull(Amount) |> scales::dollar()`) of awards, respectively. 

```{ojs}
groups_2 = transpose(agency_data)

agencyChart = {
  const width = 760, height = 460;
  const radius = Math.min(width * 0.34, height / 2 - 20);

  const svg = d3.create("svg")
      .attr("viewBox", [-width / 2, -height / 2, width, height])
      .attr("style", "max-width: 100%; height: auto; font: 13px Authentic Sans;");

  // Same endpoints as the Council Member pie, spread across however many rows
  // agency_data has. Lab interpolation keeps the lightness steps even.
  const ramp = d3.interpolateLab("#3258a3", "#c7e1f6");
  const n = groups_2.length;
  const color = d3.scaleOrdinal()
      .domain(groups_2.map(d => d.Agency))
      .range(groups_2.map((_, i) => ramp(n === 1 ? 0 : i / (n - 1))))
      .unknown("#8a8a8a");

  const arc      = d3.arc().innerRadius(0).outerRadius(radius * 0.8);
  const outerArc = d3.arc().innerRadius(radius * 0.9).outerRadius(radius * 0.9);
  const midAngle = d => {
    const a = (d.startAngle + (d.endAngle - d.startAngle) / 2) % (2 * Math.PI);
    return a < 0 ? a + 2 * Math.PI : a;
  };
  const rotate = -Math.PI / 4;
  const arcs = d3.pie()
      .sort(null)
      .startAngle(rotate)
      .endAngle(rotate + 2 * Math.PI)
      .value(d => d.pct)(groups_2);

  const labels = arcs.map(d => {
    const side = midAngle(d) < Math.PI ? 1 : -1;
    return { d, side, y: outerArc.centroid(d)[1], x: radius * side };
  });

  const minGap = 34;
  for (const side of [1, -1]) {
    const col = labels.filter(l => l.side === side).sort((a, b) => a.y - b.y);
    for (let i = 1; i < col.length; i++)
      col[i].y = Math.max(col[i].y, col[i - 1].y + minGap);
    const overflow = col.length ? col[col.length - 1].y - (height / 2 - 20) : 0;
    if (overflow > 0) col.forEach(l => l.y -= overflow);
  }

  svg.append("g")
    .selectAll("path")
    .data(labels)
    .join("path")
      .attr("d", l => arc(l.d))
      .attr("fill", l => color(l.d.data.Agency))
      .attr("stroke", "white")
      .attr("stroke-width", 2)
    .append("title")
      .text(l => `${l.d.data.Agency}: ${l.d.data.pct.toFixed(1)}%`);

  svg.append("g")
      .attr("fill", "none")
      .attr("stroke", "#666")
      .attr("stroke-width", 1)
      .attr("opacity", 0.5)
    .selectAll("polyline")
    .data(labels)
    .join("polyline")
      .attr("points", l => [
        arc.centroid(l.d),
        outerArc.centroid(l.d),
        [l.x * 0.98, l.y]
      ]);

  svg.append("g")
    .selectAll("text")
    .data(labels)
    .join("text")
      .attr("transform", l => `translate(${l.x},${l.y})`)
      .attr("text-anchor", l => l.side === 1 ? "start" : "end")
      .attr("dy", "0.35em")
      .call(t => t.append("tspan")
          .attr("font-weight", "bold")
          .text(l => l.d.data.Agency))
      .call(t => t.append("tspan")
          .attr("x", 0).attr("dy", "1.2em")
          .attr("fill", "#555")
          .text(l => `${l.d.data.pct.toFixed(1)}%`));

  return svg.node();
}
```

Interestingly, FY27 awards still list DoITT as a contracting agency, despite being renamed to OTI in a 2022 consolidation effort. 
```{r}
top_agencies_reformat <- top_agencies |> 
  mutate(Amount = scales::dollar(Amount)) |> 
  rename(
    "Total Amount" = "Amount",
    "Number of Awards" = "Number",
    "Share of Funding (%)" = "pct"
    ) 

cil_table(top_agencies_reformat, full_width = TRUE)
```

```{r}
#| message: FALSE
#| warning: FALSE
# 
# # Unload NCCS package to mitigate arrow conflict
# detach("package:nccsdata", unload = TRUE)
# library(sf)
# library(mapgl)
# 
# comDist <- read_sf("https://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/NYC_Community_Districts/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=pgeojson")
# 
# # Create spatial dataframe for schedule c
# # Use coords from BMF
# # Could also geocode schedule C address
# schedule_c_geo <- schedule_c_bmf |> 
#   filter(!is.na(geo_lon) & !is.na(geo_lat)) |> 
#   st_as_sf(coords = c("geo_lon", "geo_lat"),
#            crs = 4326) |> 
#   st_join(comDist |> select(BoroCD))
# 
# 
# # Aggregations
# comDist_all_awards <- schedule_c_geo |> 
#   st_drop_geometry() |> 
#   group_by(BoroCD) |> 
#   summarize(n_all_awards = n(),
#             dollar_all_awards = sum(as.numeric(Amount)))
# 
# comDist_arts_awards <- schedule_c_geo |> 
#   st_drop_geometry() |> 
#   filter(ntee_code_major_group == "Arts, Culture and Humanities" | Source %in% arts_culture_inits | Agency == "DCLA") |> 
#   group_by(BoroCD) |> 
#   summarize(n_arts_awards = n(),
#             dollar_arts_awards = sum(as.numeric(Amount)))
# 
# 
# comDist_agg <- comDist |> 
#   left_join(comDist_all_awards) |> 
#   left_join(comDist_arts_awards) |> 
#   select(-Shape__Length, -Shape__Area) |> 
#   mutate(across(c(n_arts_awards, dollar_arts_awards), \(x) coalesce(x, 0)),
#        pct_arts_awards  = round(n_arts_awards / n_all_awards * 100, 1),
#        pct_arts_dollars = round(dollar_arts_awards / dollar_all_awards * 100, 1))
# 

```

```{r}
#| label: comdist-choropleth

# # Sequential ramp: light -> dark maps directly to ascending values.
# pal_seq    <- c('#c7e9c0', '#a1d99b', '#74c476', '#41ab5d', '#238b45')
# col_nodata <- "#9e9e9e"
# 
# fmt_n <- scales::label_comma(accuracy = 1)
# fmt_d <- scales::label_dollar(scale_cut = scales::cut_short_scale())
# fmt_p <- scales::label_percent(scale = 1, accuracy = 0.1)
# 
# # One entry per mapped column: legend label and number format.
# layer_spec <- list(
#   n_all_awards       = list(label = "All awards (count)",    fmt = fmt_n),
#   dollar_all_awards  = list(label = "All awards (dollars)",  fmt = fmt_d),
#   n_arts_awards      = list(label = "Arts awards (count)",   fmt = fmt_n),
#   dollar_arts_awards = list(label = "Arts awards (dollars)", fmt = fmt_d),
#   pct_arts_awards    = list(label = "Arts share of awards",  fmt = fmt_p),
#   pct_arts_dollars   = list(label = "Arts share of dollars", fmt = fmt_p)
# )
# 
# # Everything to the right of geometry. Capture this BEFORE the has_/tip_ columns
# # are added below — they append to the end too, and would otherwise be mapped.
# map_cols <- names(comDist_agg)[
#   (which(names(comDist_agg) == attr(comDist_agg, "sf_column")) + 1):ncol(comDist_agg)
# ]
# stopifnot(setequal(map_cols, names(layer_spec)))
# 
# # Award counts and dollars are heavily skewed across districts, so equal
# # intervals would put nearly every district in one class.
# make_stops <- function(x, n = length(pal_seq)) {
#   q <- unname(quantile(x, probs = seq(0, 1, length.out = n), na.rm = TRUE))
#   # interpolate() needs strictly increasing values; ties collapse quantiles.
#   if (anyDuplicated(q)) seq(min(x, na.rm = TRUE), max(x, na.rm = TRUE), length.out = n) else q
# }
# 
# # has_* drives the layer filters. An explicit 0/1 flag is used rather than
# # interpolate()'s na_color, because a district with no awards has no row to join
# # and its value arrives as JSON null — which the ramp cannot evaluate, leaving
# # the district unpainted entirely (CD 101, CD 318).
# for (cl in map_cols) {
#   comDist_agg[[paste0("has_", cl)]] <- as.integer(!is.na(comDist_agg[[cl]]))
#   comDist_agg[[paste0("tip_", cl)]] <- paste0(
#     "CD ", comDist_agg$BoroCD, " — ", layer_spec[[cl]]$label, ": ",
#     ifelse(is.na(comDist_agg[[cl]]), "no data", layer_spec[[cl]]$fmt(comDist_agg[[cl]]))
#   )
# }
# 
# # Source added once and referenced by id. Passing the sf object to each
# # add_fill_layer() instead embeds the geometry once per layer.
# m <- maplibre(style = carto_style("positron"), bounds = comDist_agg) |>
#   add_source(id = "comdist", data = comDist_agg)
# 
# for (i in seq_along(map_cols)) {
#   cl    <- map_cols[i]
#   spec  <- layer_spec[[cl]]
#   stops <- make_stops(comDist_agg[[cl]])
#   vis   <- if (i == 1) "visible" else "none"
# 
#   m <- m |>
#     # Districts with no awards, drawn in a flat grey so "no data" reads as a
#     # deliberate class rather than a hole in the map.
#     add_fill_layer(
#       id     = paste0(cl, "__nodata"),
#       source = "comdist",
#       filter = list("==", list("get", paste0("has_", cl)), 0),
#       fill_color = col_nodata, fill_opacity = 1, fill_outline_color = "#ffffff",
#       tooltip    = paste0("tip_", cl),
#       visibility = vis
#     ) |>
#     add_fill_layer(
#       id     = cl,
#       source = "comdist",
#       filter = list("==", list("get", paste0("has_", cl)), 1),
#       # In interpolate(), `values` are the data breaks and `stops` are the colors.
#       fill_color = interpolate(column = cl, values = stops, stops = pal_seq),
#       fill_opacity = 1,
#       # The lightest steps sit close to the basemap, so the outline carries the
#       # district boundary rather than the fill.
#       fill_outline_color = "#ffffff",
#       tooltip       = paste0("tip_", cl),
#       hover_options = list(fill_color = "#0b0b0b", fill_opacity = 0.15),
#       visibility    = vis
#     ) |>
#     add_legend(
#       spec$label,
#       values   = spec$fmt(stops),
#       colors   = pal_seq,
#       type     = "continuous",
#       position = "bottom-left",
#       layer_id = cl,      # ties legend visibility to its layer
#       add      = i > 1
#     )
# }
# 
# m |>
#   # No layer_id, so this swatch stays up whichever measure is showing.
#   add_legend("", values = "No data", colors = col_nodata, type = "categorical",
#              position = "bottom-left", add = TRUE) |>
#   add_layers_control(
#     position = "top-right",
#     layers   = setNames(lapply(map_cols, \(cl) c(cl, paste0(cl, "__nodata"))),
#                         vapply(layer_spec[map_cols], \(s) s$label, ""))
#   ) |>
#   # mapgl's layers control toggles each entry independently, so two measures can
#   # be lit at once and the later one simply covers the earlier. Intercept in the
#   # capture phase: switch off whatever is active before the native handler turns
#   # the clicked entry on, and swallow clicks on the entry already showing.
#   htmlwidgets::onRender("
# function(el, x) {
#   var guard = false;
#   el.addEventListener('click', function(e) {
#     if (guard) return;
#     var link = e.target.closest('.layers-control .layers-list a[data-layer-ids]');
#     if (!link) return;
#     if (link.classList.contains('active')) { e.preventDefault(); e.stopPropagation(); return; }
#     guard = true;
#     el.querySelectorAll('.layers-control .layers-list a.active').forEach(function(o) { o.click(); });
#     guard = false;
#   }, true);
# }")
```
